Structure Learning for $\VAR(p)$ Models

### 2019-06-11 paper/paper.tex
Fixing "_fast" computation methods.

%% bibliography stuff -- this needs come before the preamble inclusion
\usepackage[backend=bibtex,sorting=none]{biblatex}
\usepackage{enumitem}
\usepackage{etex,etoolbox}
\usepackage{hyperref}
\usepackage{todonotes}
\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}

@@

\input{\string~/Documents/academics/global_academics/latex_preamble}

@@

We carry this out via the forward search described in Algorithm
\ref{alg:maximize_bic}.  Since $BIC(p)$ need not be a convex function,
this scheme is not guaranteed to find the true maximum of the BIC, but
this approximation suffices for our purposes \hl{TODO: is this
  sufficient?  A full search is not so much harder.}

\begin{algorithm}
  \caption{BIC Model Selection Algorithm}
  \label{alg:maximize_bic}

  \SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
  \SetKwInOut{Initialize}{initialize}
  \DontPrintSemicolon

  \BlankLine
  \Input{$T$ samples of WSS processes $x_i(t), x_j(t)$.  A maximum possible lag $p_{\text{max}}$.}
  \Output{Model orders $p_i$ and $p$ for models in Hypotheses $H_0$ and $H_1$ respectively.}
  \Initialize{
    $BIC_{\text{univariate}} = -\infty$\\
    $BIC_{\text{bivariate}} = -\infty$\\
  }

  \# \texttt{Find $p_i$}\\
  \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
    $\xi_i(p) = $\ \texttt{fit\_univariate}$(x_i(t), p)$ \texttt{ \# Fit model for $H_0$}\\
    $bic_p = -(T - p - 1) \ln \xi_i(p) - p\ln(T - p - 1)$\\
    \If{$bic_p > BIC_{\text{univariate}}$}{
      $BIC_{\text{univariate}} = bic_p$\\
      \texttt{continue}\\
    } \Else{
      $p_i = p - 1$\\
      \texttt{break}\\
    }
  }

  \BlankLine
  \# \texttt{Find $p$}\\
  \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
    $\xi_{ij}(p) = $\ \texttt{fit\_biivariate}$(x_i(t), x_j(t), p)$ \texttt{ \# Fit model for $H_1$}\\
    $bic_p = -(T - p - 1) \ln \xi_{ij}(p) - 2p\ln(T - p - 1)$\\
    \If{$bic_p > BIC_{\text{bivariate}}$}{
      $BIC_{\text{biivariate}} = bic_p$\\
      \texttt{continue}\\
    }\Else{
      $p = p - 1$\\
      \texttt{break}\\
    }
  }
    \Return{$p_i, p$}
\end{algorithm}

>>>>

%% bibliography stuff -- this needs come before the preamble inclusion
\usepackage[backend=bibtex,sorting=none]{biblatex}
\usepackage{enumitem}
\usepackage{etex,etoolbox}
\usepackage{hyperref}
\usepackage{todonotes}
\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
% \bibliography{\string~global_bib.bib}

@@

\input{\string~/Documents/academics/global_academics/latex_preamble}
% \input{\string~latex_preamble}

@@

We carry this out by a simple direct search on each model order
between $1$ and some prescribed $p_\text{max}$.  In practice it is
sufficient to pick $p_\text{max}$ ad-hoc or via some simple heuristic
(e.g. plotting the sequence $BIC(p)$ over $p$), and we will review in
Section \ref{sec:efficient_model_estimation} an algorithm to carry out
this search efficiently.

% \begin{algorithm}
%   \caption{BIC Model Selection Algorithm}
%   \label{alg:maximize_bic}

%   \SetKwInOut{Input}{input}
%   \SetKwInOut{Output}{output}
%   \SetKwInOut{Initialize}{initialize}
%   \DontPrintSemicolon

%   \BlankLine
%   \Input{$T$ samples of WSS processes $x_i(t), x_j(t)$.  A maximum possible lag $p_{\text{max}}$.}
%   \Output{Model orders $p_i$ and $p$ for models in Hypotheses $H_0$ and $H_1$ respectively.}
%   \Initialize{
%     $BIC_{\text{univariate}} = -\infty$\\
%     $BIC_{\text{bivariate}} = -\infty$\\
%   }

%   \# \texttt{Find $p_i$}\\
%   \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
%     $\xi_i(p) = $\ \texttt{fit\_univariate}$(x_i(t), p)$ \texttt{ \# Fit model for $H_0$}\\
%     $bic_p = -(T - p - 1) \ln \xi_i(p) - p\ln(T - p - 1)$\\
%     \If{$bic_p > BIC_{\text{univariate}}$}{
%       $BIC_{\text{univariate}} = bic_p$\\
%       \texttt{continue}\\
%     } \Else{
%       $p_i = p - 1$\\
%       \texttt{break}\\
%     }
%   }

%   \BlankLine
%   \# \texttt{Find $p$}\\
%   \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
%     $\xi_{ij}(p) = $\ \texttt{fit\_biivariate}$(x_i(t), x_j(t), p)$ \texttt{ \# Fit model for $H_1$}\\
%     $bic_p = -(T - p - 1) \ln \xi_{ij}(p) - 2p\ln(T - p - 1)$\\
%     \If{$bic_p > BIC_{\text{bivariate}}$}{
%       $BIC_{\text{biivariate}} = bic_p$\\
%       \texttt{continue}\\
%     }\Else{
%       $p = p - 1$\\
%       \texttt{break}\\
%     }
%   }
%     \Return{$p_i, p$}
% \end{algorithm}

\subsection{Efficient Model Estimation}
\label{sec:efficient_model_estimation}
In practice, the vast majority of computational effort involved in
implementing our estimation algorithm is spent calculating the error
estimates $\xi_i(p_i)$ and $\xi_{ij}(p, p)$.  This requires fitting a
total of $n^2p_{\text{max}}$ autoregressive models, where the most
naive algorithm (e.g. solving a least squares problem for each model)
for this task will consume $O(n^2p_{\text{max}}^4T)$ time, it is
possible to carry out this task in a much more modest
$O(n^2p_{\text{max}}^2 ) + O(n^2p_{\text{max}}T)$ time.

Firstly, we estimate the sequence of Covariance
matrices\footnote{The particular indexing and normalization given in
  equation \ref{eqn:covariance_estimate} is critical to ensure
  $\widehat{R}$ is positive semidefinite.  The estimate can be viewed
  as calculating the covariance sequence of a signal multiplied by a
  rectangular window.}

\begin{equation}
  \label{eqn:covariance_estimate}
  \widehat{R}_x(\tau) = \frac{1}{T}\sum_{t = \tau + 1}^T x(t) x(t - \tau)^\T;\ \tau = 0, \ldots, p_{\text{max}},
\end{equation}

an $O(n^2p_{\text{max}}T)$ operation.  Secondly, the variance
estimates $\xi_i(p)$ for $p = 1, \ldots, p_{\text{max}}$ can be
evaluated in $O(p_{\text{max}}^2)$ time each by applying the
Levinson-Durbin recursion to $\widehat{R}_{ii}(tau)$, which
effectively estimates a sequence of $AR$ models, producing $\xi_i(p)$
as a side-effect (see
\cite{hayes_statistical_digital_signal_processing} and
\cite{levinson_durbin_recursion}).

Similarly, the variance estimates $\xi_{ij}$ and $\xi_{ji}$ can be
obtained by estimating $\frac{(n + 1)n}{2}$ bivariate AR models, again
in $O(p_{\text{max}}^2)$ via Whittle's generalized LD recursion
\cite{whittle_generalized_levinson_durbin}.
